---
title: "Lab 06 - Sad plots"
author: "John Bennett"
date: "Feb. 22, 2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r Housekeeping}
library(usethis)
use_git_config(user.name = "John Bennett", 
               user.email = "bennjt20@wfu.edu")
```

```{r Housekeeping2}
library(tidyverse)
```

```{r Staffcsv}
staff <- read_csv("data/instructional-staff.csv")
```

```{r pivot}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```

```{r lineplot}
staff_long %>%
  ggplot(aes(x = year, y = value, color = faculty_type)) +
  geom_line()
```

```{r lineplot}
staff_long %>%
  ggplot(aes(x = year, y = value, group = faculty_type, color = faculty_type)) +
  geom_line() +
labs(x = "Years for which we have data",
   y = "Percentage of hires, by type of faculty for each year",
   title = "Instructional staff employees between 1975 and 2011",
   color = "Faculty Type"
 )
```



### Exercise 1
Note: See Lineplot.png

Q: a brief paragraph describing the choices you made in your improvement, specifically discussing what you didn’t like in the original plots and why, and how you addressed them in the visualization you created.
A: My goal is to visualize the most relevant aspect of the data. The way to do that is to show how the proportions changed over time. In other words, to show how one year was different than another. The best way to visualize that would be to show proportions for the individual years. That way, there is a solid block of color representing the Faculty Type, and the solid blocks are of visibly different size when there are different proportions from one year to another. The current line graph (geom_line) display does not capture this fundamental quantitative material. The lines obscure the quantified proportion we're interested in, which is a "block" of data and should be represented by a block, I think.

### Exercise 2
Q: Suppose the objective of this plot was to show that the proportion of part-time faculty have gone up over time compared to other instructional staff types. What changes would you propose making to this plot to tell this story?

A: I would use (position="fill") to create a digestible visual showing the percentage of hires by year. Within each year bar, there would be five different colors for each of the faculty types. Some of my proposed codes are below.

```{r FacetedBarPlot1}
ggplot(staff_long, aes(y = year, x = value, fill = faculty_type)) +
  geom_col(position = "fill")
```



### Exercise 3
My first bit of advice would be to chop the top 3/4 off of the harvested tons visual. Above 20 million tons, there is no useful information. Unless one wants to highlight China's vast harvest, there really is no point to the upper part of the visual. 
Additionally, the aggregate data can be interesting, but also unhelpful for understanding a country's per capita harvest. If we want to look at a society's consumption level, individual-level measurements make sense as a gauge that can be compared with other nations' per capita averages. 
Also, if our friend cannot articulate a compelling reason for having separate visuals for aquaculture and capture, I would recommend that he combined them in order to de-clutter the visualization. 

```{r fisheries}
fisheries <- read_csv("data/fisheries.csv")               
```


#Stretch Practice

```{r fisheries}
library(tidyverse) 
library(mosaicData)               
```

```{r fisheries}
data(Whickham)  
?Whickham
```


Q1.What type of study do you think these data come from: observational or experiment? Why?
A: Observational, because there was no random assignment

Q2. How many observations are in this dataset? What does each observation represent?
A: 1314 observations on women on variables pertaining to smoking and mortality.

Q3. How many variables are in this dataset? What type of variable is each? Display each variable using an appropriate visualization.
A: 3 variables in the dataset: 
1) outcome survival status after 20 years: a factor with levels Alive Dead
2) smoker smoking status at baseline: a factor with levels No Yes
3) age (in years) at the time of the first survey

Q4. What would you expect the relationship between smoking status and health outcome to be?
A: I would expect smoking to be associated with substantially lower survival outcomes.

Q5. Create a visualization depicting the relationship between smoking status and health outcome. Briefly describe the relationship, and evaluate whether this meets your expectations. Additionally, calculate the relevant conditional probabilities to help your narrative. 
A: 
P(Alive|non-Smoker) = 0.69
P(Dead|non-Smoker) = 0.31
P(Alive|Smoker) = 0.76
P(Dead|Smoker) = 0.24


```{r smoking}
Whickham %>%
  count(smoker, outcome)
```

```{r smokers1}
Whickham %>%
  count(smoker, outcome) %>%
  group_by(smoker) %>%
  mutate(prop_outcome = n / sum(n))
```  

```{r smokers}
ggplot(Whickham, aes(x = outcome,
                 fill = smoker)) +     
geom_bar() +
labs(title = "Smoking is bad")
```


### Exercise 6
Create a new variable called age_cat using the following scheme:

```{r Ex6}
Whickham <- Whickham %>%
  mutate(age_cat = case_when(age <= 44 ~ "18-44",
    age > 44 & age <= 64 ~ "45-64",
    age > 64 ~ "65+"))
```

### Exercise 7
Re-create the visualization depicting the relationship between smoking status and health outcome, faceted by age_cat. What changed? What might explain this change? Extend the contingency table from earlier by breaking it down by age category and use it to help your narrative.

```{r Ex7.1}
Whickham %>%
mutate(age_cat, smoker, outcome)
```
  
  
```{r Ex7.2}
Whickham %>%
  count(age_cat, smoker, outcome) %>%
  pivot_wider(names_from = age_cat, values_from = n) 
```  
  
  scale_x_continuous(labels = label_percent(scale = 100,
  prefix = "",
  suffix = "%",
  big.mark = " ",
  decimal.mark = ".")) +
  
```{r Ex7.3}
ggplot(Whickham, aes(y = smoker, fill = outcome)) +
  geom_bar(position = "fill") +
  facet_wrap(. ~ age_cat) +
  labs(title = "Life outcome by smoking status and age group",
       x = NULL, y = "Smoker", fill = NULL) +
  theme(legend.position = "bottom")
```   
  
  
  It appears that the code worked, however there is no meaninful output that I am aware of. 
  
```{r Ex7.4}
  Whickham %>%
  count(age_cat, smoker, outcome) %>%
  group_by(age_cat, smoker) %>%
  mutate(
    n_smoker  = sum(n),
    prop_fatality = n / n_smoker
    ) %>%
  filter(outcome == "Fatality") %>%
  rename(n_fatality = n) %>%
  select(-outcome) %>%
  print(n = 20)
```
  




  

